---
title: "Using Machine Learning Algorithms for the Computer-Aided Assessment of Periapical Lesions in 2D Periapical Radiography Images"
author: "C. Bennasar$^1$, Y. Cid$^{1,2}$, I. García$^{1,2,*}$, J. Jiménez$^1$ and F. Pérez$^3$"
date: "09/04/2020"
output: html_document
---
```{r echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE}
library (MASS)
library (ISLR)
library (leaps)
library(ggstatsplot)
library(ggplot2)
library(caret)
library(car)
library(psych)
library(randomForest)
require(nnet)
require(reshape2)
library(class)
library(e1071)
library(neuralnet)
library(irr)
library( epiR)
```

```{r echo=FALSE}
library(readxl)
datos <- read_excel("datos.xlsx")
datos.v=datos[c("cat","rob","ram","var","per","verg","balle")]

N=500
s=1.0
k2.cat=rep(0,N)
k3.cat=rep(0,N)
k4.cat=rep(0,N)
k5.cat=rep(0,N)
k6.cat=rep(0,N)
k7.cat=rep(0,N)

k3.rob=rep(0,N)
k4.rob=rep(0,N)
k5.rob=rep(0,N)
k6.rob=rep(0,N)
k7.rob=rep(0,N)

k4.ram=rep(0,N)
k5.ram=rep(0,N)
k6.ram=rep(0,N)
k7.ram=rep(0,N)

k5.var=rep(0,N)
k6.var=rep(0,N)
k7.var=rep(0,N)

k6.per=rep(0,N)
k7.per=rep(0,N)

k7.verg=rep(0,N)

for(i in 1:N)
{
                #set.seed(100)
                index = sample(seq_len(nrow(datos.v)), size = s*nrow(datos.v), replace=T)
                boostraped = datos.v[index,]
                
                cat2=boostraped[c(1,2)]
                k2.cat[i]=kappam.fleiss(cat2, detail = T)$value
                cat3=boostraped[c(1,3)]
                k3.cat[i]=kappam.fleiss(cat3, detail = T)$value
                cat4=boostraped[c(1,4)]
                k4.cat[i]=kappam.fleiss(cat4, detail = T)$value
                cat5=boostraped[c(1,5)]
                k5.cat[i]=kappam.fleiss(cat5, detail = T)$value
                cat6=boostraped[c(1,6)]
                k6.cat[i]=kappam.fleiss(cat6, detail = T)$value
                cat7=boostraped[c(1,7)]
                k7.cat[i]=kappam.fleiss(cat7, detail = T)$value
                
                rob3=boostraped[c(2,3)]
                k3.rob[i]=kappam.fleiss(rob3, detail = T)$value
                rob4=boostraped[c(2,4)]
                k4.rob[i]=kappam.fleiss(rob4, detail = T)$value
                rob5=boostraped[c(2,5)]
                k5.rob[i]=kappam.fleiss(rob5, detail = T)$value
                rob6=boostraped[c(2,6)]
                k6.rob[i]=kappam.fleiss(rob6, detail = T)$value
                rob7=boostraped[c(2,7)]
                k7.rob[i]=kappam.fleiss(rob7, detail = T)$value
                
                ram4=boostraped[c(3,4)]
                k4.ram[i]=kappam.fleiss(ram4, detail = T)$value
                ram5=boostraped[c(3,5)]
                k5.ram[i]=kappam.fleiss(ram5, detail = T)$value
                ram6=boostraped[c(3,6)]
                k6.ram[i]=kappam.fleiss(ram6, detail = T)$value
                ram7=boostraped[c(3,7)]
                k7.ram[i]=kappam.fleiss(ram7, detail = T)$value
                
                var5=boostraped[c(4,5)]
                k5.var[i]=kappam.fleiss(var5, detail = T)$value
                var6=boostraped[c(4,6)]
                k6.var[i]=kappam.fleiss(var6, detail = T)$value
                var7=boostraped[c(4,7)]
                k7.var[i]=kappam.fleiss(var7, detail = T)$value
                
                per6=boostraped[c(5,6)]
                k6.per[i]=kappam.fleiss(per6, detail = T)$value
                per7=boostraped[c(5,7)]
                k7.per[i]=kappam.fleiss(per7, detail = T)$value
                
                verg7=boostraped[c(6,7)]
                k7.verg[i]=kappam.fleiss(verg7, detail = T)$value
               
}
k.v=c(k2.cat,k3.cat,k4.cat,k5.cat,k6.cat,k7.cat,k3.rob,k4.rob,k5.rob,k6.rob,k7.rob,k4.ram,k5.ram,k6.ram,k7.ram,k5.var,k6.var,k7.var,k6.per,k7.per,k7.verg)
hist(k.v)
summary(k.v)
quantile(k.v,0.025)

library(readxl)
datos.tot <- read_excel("datos.xlsx")
datos=datos.tot[datos.tot$position=="IC",]
datos.v=datos[c("cat","rob","ram","var","per","verg","balle")]

N=500
s=1.0
k2.cat=rep(0,N)
k3.cat=rep(0,N)
k4.cat=rep(0,N)
k5.cat=rep(0,N)
k6.cat=rep(0,N)
k7.cat=rep(0,N)

k3.rob=rep(0,N)
k4.rob=rep(0,N)
k5.rob=rep(0,N)
k6.rob=rep(0,N)
k7.rob=rep(0,N)

k4.ram=rep(0,N)
k5.ram=rep(0,N)
k6.ram=rep(0,N)
k7.ram=rep(0,N)

k5.var=rep(0,N)
k6.var=rep(0,N)
k7.var=rep(0,N)

k6.per=rep(0,N)
k7.per=rep(0,N)

k7.verg=rep(0,N)

for(i in 1:N)
{
                #set.seed(100)
                index = sample(seq_len(nrow(datos.v)), size = s*nrow(datos.v), replace=T)
                boostraped = datos.v[index,]
                
                cat2=boostraped[c(1,2)]
                k2.cat[i]=kappam.fleiss(cat2, detail = T)$value
                cat3=boostraped[c(1,3)]
                k3.cat[i]=kappam.fleiss(cat3, detail = T)$value
                cat4=boostraped[c(1,4)]
                k4.cat[i]=kappam.fleiss(cat4, detail = T)$value
                cat5=boostraped[c(1,5)]
                k5.cat[i]=kappam.fleiss(cat5, detail = T)$value
                cat6=boostraped[c(1,6)]
                k6.cat[i]=kappam.fleiss(cat6, detail = T)$value
                cat7=boostraped[c(1,7)]
                k7.cat[i]=kappam.fleiss(cat7, detail = T)$value
                
                rob3=boostraped[c(2,3)]
                k3.rob[i]=kappam.fleiss(rob3, detail = T)$value
                rob4=boostraped[c(2,4)]
                k4.rob[i]=kappam.fleiss(rob4, detail = T)$value
                rob5=boostraped[c(2,5)]
                k5.rob[i]=kappam.fleiss(rob5, detail = T)$value
                rob6=boostraped[c(2,6)]
                k6.rob[i]=kappam.fleiss(rob6, detail = T)$value
                rob7=boostraped[c(2,7)]
                k7.rob[i]=kappam.fleiss(rob7, detail = T)$value
                
                ram4=boostraped[c(3,4)]
                k4.ram[i]=kappam.fleiss(ram4, detail = T)$value
                ram5=boostraped[c(3,5)]
                k5.ram[i]=kappam.fleiss(ram5, detail = T)$value
                ram6=boostraped[c(3,6)]
                k6.ram[i]=kappam.fleiss(ram6, detail = T)$value
                ram7=boostraped[c(3,7)]
                k7.ram[i]=kappam.fleiss(ram7, detail = T)$value
                
                var5=boostraped[c(4,5)]
                k5.var[i]=kappam.fleiss(var5, detail = T)$value
                var6=boostraped[c(4,6)]
                k6.var[i]=kappam.fleiss(var6, detail = T)$value
                var7=boostraped[c(4,7)]
                k7.var[i]=kappam.fleiss(var7, detail = T)$value
                
                per6=boostraped[c(5,6)]
                k6.per[i]=kappam.fleiss(per6, detail = T)$value
                per7=boostraped[c(5,7)]
                k7.per[i]=kappam.fleiss(per7, detail = T)$value
                
                verg7=boostraped[c(6,7)]
                k7.verg[i]=kappam.fleiss(verg7, detail = T)$value
               
}
k.v=c(k2.cat,k3.cat,k4.cat,k5.cat,k6.cat,k7.cat,k3.rob,k4.rob,k5.rob,k6.rob,k7.rob,k4.ram,k5.ram,k6.ram,k7.ram,k5.var,k6.var,k7.var,k6.per,k7.per,k7.verg)
hist(k.v)
summary(k.v)
quantile(k.v,0.025)
quantile(k.v,0.975)

library(readxl)
datos.tot <- read_excel("datos.xlsx")
datos=datos.tot[datos.tot$position=="PM",]
datos.v=datos[c("cat","rob","ram","var","per","verg","balle")]

N=500
s=1.0
k2.cat=rep(0,N)
k3.cat=rep(0,N)
k4.cat=rep(0,N)
k5.cat=rep(0,N)
k6.cat=rep(0,N)
k7.cat=rep(0,N)

k3.rob=rep(0,N)
k4.rob=rep(0,N)
k5.rob=rep(0,N)
k6.rob=rep(0,N)
k7.rob=rep(0,N)

k4.ram=rep(0,N)
k5.ram=rep(0,N)
k6.ram=rep(0,N)
k7.ram=rep(0,N)

k5.var=rep(0,N)
k6.var=rep(0,N)
k7.var=rep(0,N)

k6.per=rep(0,N)
k7.per=rep(0,N)

k7.verg=rep(0,N)

for(i in 1:N)
{
                #set.seed(100)
                index = sample(seq_len(nrow(datos.v)), size = s*nrow(datos.v), replace=T)
                boostraped = datos.v[index,]
                
                cat2=boostraped[c(1,2)]
                k2.cat[i]=kappam.fleiss(cat2, detail = T)$value
                cat3=boostraped[c(1,3)]
                k3.cat[i]=kappam.fleiss(cat3, detail = T)$value
                cat4=boostraped[c(1,4)]
                k4.cat[i]=kappam.fleiss(cat4, detail = T)$value
                cat5=boostraped[c(1,5)]
                k5.cat[i]=kappam.fleiss(cat5, detail = T)$value
                cat6=boostraped[c(1,6)]
                k6.cat[i]=kappam.fleiss(cat6, detail = T)$value
                cat7=boostraped[c(1,7)]
                k7.cat[i]=kappam.fleiss(cat7, detail = T)$value
                
                rob3=boostraped[c(2,3)]
                k3.rob[i]=kappam.fleiss(rob3, detail = T)$value
                rob4=boostraped[c(2,4)]
                k4.rob[i]=kappam.fleiss(rob4, detail = T)$value
                rob5=boostraped[c(2,5)]
                k5.rob[i]=kappam.fleiss(rob5, detail = T)$value
                rob6=boostraped[c(2,6)]
                k6.rob[i]=kappam.fleiss(rob6, detail = T)$value
                rob7=boostraped[c(2,7)]
                k7.rob[i]=kappam.fleiss(rob7, detail = T)$value
                
                ram4=boostraped[c(3,4)]
                k4.ram[i]=kappam.fleiss(ram4, detail = T)$value
                ram5=boostraped[c(3,5)]
                k5.ram[i]=kappam.fleiss(ram5, detail = T)$value
                ram6=boostraped[c(3,6)]
                k6.ram[i]=kappam.fleiss(ram6, detail = T)$value
                ram7=boostraped[c(3,7)]
                k7.ram[i]=kappam.fleiss(ram7, detail = T)$value
                
                var5=boostraped[c(4,5)]
                k5.var[i]=kappam.fleiss(var5, detail = T)$value
                var6=boostraped[c(4,6)]
                k6.var[i]=kappam.fleiss(var6, detail = T)$value
                var7=boostraped[c(4,7)]
                k7.var[i]=kappam.fleiss(var7, detail = T)$value
                
                per6=boostraped[c(5,6)]
                k6.per[i]=kappam.fleiss(per6, detail = T)$value
                per7=boostraped[c(5,7)]
                k7.per[i]=kappam.fleiss(per7, detail = T)$value
                
                verg7=boostraped[c(6,7)]
                k7.verg[i]=kappam.fleiss(verg7, detail = T)$value
               
}
k.v=c(k2.cat,k3.cat,k4.cat,k5.cat,k6.cat,k7.cat,k3.rob,k4.rob,k5.rob,k6.rob,k7.rob,k4.ram,k5.ram,k6.ram,k7.ram,k5.var,k6.var,k7.var,k6.per,k7.per,k7.verg)
hist(k.v)
summary(k.v)
quantile(k.v,0.025)
quantile(k.v,0.975)

library(readxl)
datos.tot <- read_excel("datos.xlsx")
datos=datos.tot[datos.tot$position=="M",]
datos.v=datos[c("cat","rob","ram","var","per","verg","balle")]

N=500
s=1.0
k2.cat=rep(0,N)
k3.cat=rep(0,N)
k4.cat=rep(0,N)
k5.cat=rep(0,N)
k6.cat=rep(0,N)
k7.cat=rep(0,N)

k3.rob=rep(0,N)
k4.rob=rep(0,N)
k5.rob=rep(0,N)
k6.rob=rep(0,N)
k7.rob=rep(0,N)

k4.ram=rep(0,N)
k5.ram=rep(0,N)
k6.ram=rep(0,N)
k7.ram=rep(0,N)

k5.var=rep(0,N)
k6.var=rep(0,N)
k7.var=rep(0,N)

k6.per=rep(0,N)
k7.per=rep(0,N)

k7.verg=rep(0,N)

for(i in 1:N)
{
                #set.seed(100)
                index = sample(seq_len(nrow(datos.v)), size = s*nrow(datos.v), replace=T)
                boostraped = datos.v[index,]
                
                cat2=boostraped[c(1,2)]
                k2.cat[i]=kappam.fleiss(cat2, detail = T)$value
                cat3=boostraped[c(1,3)]
                k3.cat[i]=kappam.fleiss(cat3, detail = T)$value
                cat4=boostraped[c(1,4)]
                k4.cat[i]=kappam.fleiss(cat4, detail = T)$value
                cat5=boostraped[c(1,5)]
                k5.cat[i]=kappam.fleiss(cat5, detail = T)$value
                cat6=boostraped[c(1,6)]
                k6.cat[i]=kappam.fleiss(cat6, detail = T)$value
                cat7=boostraped[c(1,7)]
                k7.cat[i]=kappam.fleiss(cat7, detail = T)$value
                
                rob3=boostraped[c(2,3)]
                k3.rob[i]=kappam.fleiss(rob3, detail = T)$value
                rob4=boostraped[c(2,4)]
                k4.rob[i]=kappam.fleiss(rob4, detail = T)$value
                rob5=boostraped[c(2,5)]
                k5.rob[i]=kappam.fleiss(rob5, detail = T)$value
                rob6=boostraped[c(2,6)]
                k6.rob[i]=kappam.fleiss(rob6, detail = T)$value
                rob7=boostraped[c(2,7)]
                k7.rob[i]=kappam.fleiss(rob7, detail = T)$value
                
                ram4=boostraped[c(3,4)]
                k4.ram[i]=kappam.fleiss(ram4, detail = T)$value
                ram5=boostraped[c(3,5)]
                k5.ram[i]=kappam.fleiss(ram5, detail = T)$value
                ram6=boostraped[c(3,6)]
                k6.ram[i]=kappam.fleiss(ram6, detail = T)$value
                ram7=boostraped[c(3,7)]
                k7.ram[i]=kappam.fleiss(ram7, detail = T)$value
                
                var5=boostraped[c(4,5)]
                k5.var[i]=kappam.fleiss(var5, detail = T)$value
                var6=boostraped[c(4,6)]
                k6.var[i]=kappam.fleiss(var6, detail = T)$value
                var7=boostraped[c(4,7)]
                k7.var[i]=kappam.fleiss(var7, detail = T)$value
                
                per6=boostraped[c(5,6)]
                k6.per[i]=kappam.fleiss(per6, detail = T)$value
                per7=boostraped[c(5,7)]
                k7.per[i]=kappam.fleiss(per7, detail = T)$value
                
                verg7=boostraped[c(6,7)]
                k7.verg[i]=kappam.fleiss(verg7, detail = T)$value
               
}
k.v=c(k2.cat,k3.cat,k4.cat,k5.cat,k6.cat,k7.cat,k3.rob,k4.rob,k5.rob,k6.rob,k7.rob,k4.ram,k5.ram,k6.ram,k7.ram,k5.var,k6.var,k7.var,k6.per,k7.per,k7.verg)
hist(k.v)
summary(k.v)
quantile(k.v,0.025)
quantile(k.v,0.975)

library(readxl)
datos <- read_excel("datos.xlsx")
datos$datos1=round(datos$cat+datos$per+datos$balle/3)
datos$datos2=round(datos$rob+datos$var+datos$ram/3)

N=500
s=1.0
k.g=rep(0,N)
for(i in 1:N)
{
                #set.seed(100)
                index = sample(seq_len(nrow(datos)), size = s*nrow(datos), replace=T)
                boostraped = datos[index,]
                expertos=boostraped[c(10,11)]
                k.g[i]=kappam.fleiss(expertos, detail = T)$value
}
hist(k.g)
summary(k.g)
quantile(k.g,0.025)
quantile(k.g,0.975)
```


```{r}
ML=read.table(file = "ML.txt", header = T)
cat.dat=ML[c("position","cat.v","cat.b","cat.l")]
names(cat.dat)=c("position","v","b","l")

rob.dat=ML[c("position","rob.v","rob.b","rob.l")]
names(rob.dat)=c("position","v","b","l")

ram.dat=ML[c("position","ram.v","ram.b","ram.l")]
names(ram.dat)=c("position","v","b","l")

var.dat=ML[c("position","var.v","var.b","var.l")]
names(var.dat)=c("position","v","b","l")

per.dat=ML[c("position","per.v","per.b","per.l")]
names(per.dat)=c("position","v","b","l")

verg.dat=ML[c("position","verg.v","verg.b","verg.l")]
names(verg.dat)=c("position","v","b","l")

balle.dat=ML[c("position","balle.v","balle.b","balle.l")]
names(balle.dat)=c("position","v","b","l")

catalogo.todos=rbind(cat.dat,rob.dat,ram.dat,var.dat,verg.dat,verg.dat,balle.dat)

```


```{r warning=FALSE}
                data.lr=catalogo.todos
              
                #Random sampling
                samplesize = 0.8*nrow(data.lr)
                set.seed(100) # for reproducibility
                index = sample(seq_len(nrow(data.lr)), size = samplesize, replace = F)
                train= data.lr[index, ]
                test = data.lr[-index,]
              
                
                  # Fitting Multiple Logistic Regression and predicting
                  glm.fit= glm(v ~ ., data = data.lr, family = binomial)
                  glm.pred = predict(glm.fit, newdata = test, type = "response")
                  
                  
```


Let's see how important is each covariate in the regression:


```{r}
summary(glm.fit)
```

Next, to assess the performance of the model, I predict on the testing set, generating the confusion matrix, and computing the sensitivity and specificity:

```{r}
                  predicted = ifelse(glm.pred > 0.5,"1", "0")
                  observed= test$v
                  M=table (predicted,observed)
                  M
                  sensitivity(M)
                  specificity(M)
```

Finally the accuracy of the model (i.e. the percentage of observations that were correctly predicted in the testing set) is given by:
```{r}
A=100*(M[1,1]+M[2,2])/(length(test$v))
A
```

#### **Random Forest**

```{r warning=FALSE}
                data.rf=catalogo.todos
                data.rf$v=as.factor(data.rf$v)
                #Random sampling
                samplesize = 0.8*nrow(data.rf)
                set.seed(100) # for reproducibility
                index = sample(seq_len(nrow(data.rf)), size = samplesize, replace = F)
                train= data.rf[index, ]
                test = data.rf[-index,]
              
                
                  # Fitting Random Forest and predicting
                  rf.fit = randomForest(v ~., ntree = 500, mtry = 2, data = train)
                  importance (rf.fit)
```
The confusion matrix, sensitivity, specificity, and accuracy of this model when predicting on the testing set are given by:

```{r}
                  predicted=as.numeric(predict(rf.fit,test))-1
                  observed= test$v
                  M=table (predicted,observed)
                  M
                  sensitivity(M)
                  specificity(M)
                  A=100*(M[1,1]+M[2,2])/(length(test$v))
                  A

```

#### **Naive-Bayes:**


```{r}
ML=read.table(file = "ML.txt", header = T)
cat.dat=ML[c("position","cat.v","cat.b","cat.l")]
names(cat.dat)=c("position","v","b","l")

rob.dat=ML[c("position","rob.v","rob.b","rob.l")]
names(rob.dat)=c("position","v","b","l")

ram.dat=ML[c("position","ram.v","ram.b","ram.l")]
names(ram.dat)=c("position","v","b","l")

var.dat=ML[c("position","var.v","var.b","var.l")]
names(var.dat)=c("position","v","b","l")

per.dat=ML[c("position","per.v","per.b","per.l")]
names(per.dat)=c("position","v","b","l")

verg.dat=ML[c("position","verg.v","verg.b","verg.l")]
names(verg.dat)=c("position","v","b","l")

balle.dat=ML[c("position","balle.v","balle.b","balle.l")]
names(balle.dat)=c("position","v","b","l")

catalogo.nb=rbind(cat.dat,rob.dat,ram.dat,var.dat,verg.dat,verg.dat,balle.dat)



```



```{r warning=FALSE}
                data.nb=catalogo.nb
                data.nb$v=as.factor(data.nb$v)
                #Random sampling
                samplesize = 0.8*nrow(data.nb)
                set.seed(100) # for reproducibility
                index = sample(seq_len(nrow(data.nb)), size = samplesize, replace = F)
                train= data.nb[index, ]
                test = data.nb[-index,]
              
                
                  # Fitting Naive-Bayes and predicting
                  NBmod = naiveBayes(v ~ .,data = train)
                  NB.pred = predict(NBmod, newdata=test, type = "class")
                  predicted=as.numeric(NB.pred)-1
                  observed=test$v
        
                  M=table (predicted,observed)
                  M
                  sensitivity(M)
                  specificity(M)
                  A=100*(M[1,1]+M[2,2])/(length(test$v))
                  A
```
